---
title: "Create Floodplain Forest Predictors"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

[IWR Planning Assistance Library](https://ecolibrary.planusace.us/)

  -  Log in
  -  Search: "UMRS Floodplain Forest"
  -  Download: 
      -  "UMRS Floodplain Forest Model Documentation and Application Guide"
      -  "UMRS Floodplain Forest Model Application Spreadsheet"


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(nybem)
library(readr)
```



# Forest Surveys
The USACE [Forest Management Geodatabase](https://github.com/ForestManagementGeodatabase) project maintains a set of forest stand surveys completed on the Upper Mississippi River System (UMRS). These stand summaries can be used to create the input for the Floodplain Forest model. 

## Steps
  -  Open the [Forest Stand Summaries (opens in another window)](https://forestmanagementgeodatabase.github.io/FMG-MVR-Reports/){target="_blank"}. 
  -  Arrange your browser windows so that this window and the forest stand summaries window are side-by-side. 
  -  Locate the forest stand summaries that cover the project study area. 
  

# Import Observations
Import an existing table of Floodplain Forest model predictor variables. 
```{r}
forest_observations <- read_csv("input/forest_observations.csv",
                                show_col_types = TRUE)
```

```{r echo=FALSE, eval=FALSE}
library(DataEditR)
# https://cran.r-project.org/web/packages/DataEditR/vignettes/DataEditR.html

data_edit(forest_observations, 
          theme = "sandstone",
          col_edit = FALSE, row_edit = TRUE,
          col_names = FALSE, col_readonly = c("unit"),
          save_as = "input/forest_observations.csv", 
          read_fun = "read_csv", write_fun = "write_csv",
          hide = TRUE)
```
